Unifying the communicable disease spreading paradigm with Gompertzian growth

### 2022-12-19 arxiv/main.tex
updating intro with thea's new refs

This global biosphere disturbance could be in the form of short-term solar and geomagnetic disturbances, which have been associated with physiological changes in cardiac system processes \cite{vieira2022geomagnetic,Alabdulgader2018,ghione1998geomagnetic,chernouss2001geophysical,stoupel1995ambulatory,stoupel2011twenty,malin1979correlation,stoupel1993sudden}, melatonin/serotonin balance \cite{burch1999geomagnetic,rapoport1997effects,bergiannaki1996seasonal}, endothelial activation and inflammation \cite{schiff2022role}, and calcium fluxes \cite{blackman1990elf}, all with the ability to cause neurological and pulmonary adverse effects. More general effects have also been studied in this context, e.g. overall mortality, suicide, mental disorders, homicides, and epileptic seizures \cite{zenchenko2021possible,zilli2019geomagnetic,gordon2003effect,kay1994geomagnetic,kay2004schizophrenia,halberg2005chronomics,berk2006ambient,doronin1998effect,persinger1995sudden}. Cherry (2002) put forward a mechanism whereby the ionosphere acts as a coupling agent between geomagnetic disturbances and physiological effects based on the disturbance of the background electromagnetic oscillations between the surface of the Earth and the lower ionosphere, called the Schumann Resonance signal \cite{cherry2002schumann,schumann1952strahlungslosen}. The Schumann Resonance would then couple with the brain to alter calcium fluxes and the melatonin/serotonin balance, which was a causal link observed by Wever (1974) through the circadian rhythm \cite{wever1974elf}.

While the Schumann Resonance mechanism is postulated to be caused by natural geomagnetic disturbances, others have looked at possible causative links through man-made radiation stressors and have found a similar pattern in the symptomology of a Coronavirus infection and radiation injury \cite{rios2021commonalities,Rubik2021}. It is further conceivable to have an anthropogenic disturbance of the ionosphere through the use of e.g. satellite communication technologies \cite{siingh2005review}. Zaporozhan and Ponomarenko (2010) proposed the cryptochrome protein-family to be the designated geomagentic receptors able to pick up minute disturbances in magnetic fields which could translate into influenza symptoms \cite{zaporozhan2010mechanisms}. The cryptochrome proteins have also been linked to circadian regulation \cite{rodgers2009chemical} and have been extensively studied in the context of avian navigation \cite{hiscock2016quantum}.

>>>>

This global biosphere disturbance could be in the form of sudden ionospheric disturbances (SIDs) \cite{cannon1982schumann}, which in the context of short-term solar and geomagnetic disturbances have been associated with physiological changes in cardiac system processes \cite{vieira2022geomagnetic,Alabdulgader2018,ghione1998geomagnetic,chernouss2001geophysical,stoupel1995ambulatory,stoupel2011twenty,malin1979correlation,stoupel1993sudden}, melatonin/serotonin balance \cite{burch1999geomagnetic,rapoport1997effects,bergiannaki1996seasonal}, endothelial activation and inflammation \cite{schiff2022role}, and disturbed magnetoreception in migrating animals \cite{keeton1974normal,schreiber1976correlation,kowalski1988normal,bianco2019magnetic,ferrari2017cetacean,vanselow2018solar,granger2020gray}. More general effects have also been studied in this context, e.g. overall mortality, suicide, mental disorders, homicides, and epileptic seizures \cite{zenchenko2021possible,zilli2019geomagnetic,gordon2003effect,kay1994geomagnetic,kay2004schizophrenia,halberg2005chronomics,berk2006ambient,doronin1998effect,persinger1995sudden}. Cherry (2002) put forward a mechanism whereby the ionosphere acts as a coupling agent between geomagnetic disturbances and physiological effects based on the disturbance of the background electromagnetic oscillations between the surface of the Earth and the lower ionosphere, called the Schumann Resonance signal \cite{cherry2002schumann,schumann1952strahlungslosen}. The Schumann Resonance would then couple with the brain to alter the melatonin/serotonin balance, which was a causal link observed by Wever (1974) through the circadian rhythm \cite{wever1974elf}. Indeed, more recent findings support this hypothesized synchronization between Schumann Resonances and the brain \cite{pobachenko2006contingency,saroka2014quantitative,persinger2015human,konig2012biologic}.

In support of Cherry's theory, specific phenomena such as solar X-ray bursts \cite{dyrda2015novel,roldugin2004schumann,satori2005response,satori2016effects,shvets2017effect}, solar proton events \cite{roldugin2004schumann,schlegel1999schumann,singh2014x} and geomagnetic storms \cite{pazos2019analysis,salinas2016solar} can modify the shape of the ionosphere cavity and shift the Schumann resonance frequencies \cite{kudintseva2018modifications,satori2016effects,zhou2015studies}. Seismic activity has also been connected to variations in Schumann resonances \cite{christofilakis2019significant,galuk2019scattering,hayakawa2020scattering}, and conversely, ionospheric modifications have been connected to global seismicity \cite{Marchitelli2020}.

While Schumann Resonance disturbances can be caused by natural geomagnetic events, others have looked at possible causative links through man-made influences on the ionosphere and magnetosphere through the use of electromagnetic infrastructure \cite{siingh2005review}. It is thus conceivable to have an anthropogenic disturbance of the ionosphere through the use of e.g. satellite communication technologies. Not only might this effect the
global electric circuit \cite{rycroft2008overview} and the Schumann Resonance, but the usage of satellite communication technologies might also alter the magnetosphere itself leaving a signal similar to Power Line Harmonic Radiation \cite{bullough2017power,kositsky2001influence}.

Zaporozhan and Ponomarenko (2010) proposed the cryptochrome protein-family to be the designated geomagnetic receptors able to pick up minute disturbances in magnetic fields which could translate into influenza symptoms \cite{zaporozhan2010mechanisms}. The cryptochrome proteins have also been linked to circadian regulation \cite{rodgers2009chemical} and have been extensively studied in the context of avian navigation \cite{hiscock2016quantum}. Besides the cryptochromes, Krylov (2017) presents evidence and references that another iron-containing protein called the magnetoreceptor protein (MagR) encoded by the gene CG8198 is of equal importance \cite{krylov2017biological}. He suggested that geomagnetic storms could be perceived by animals as a disruption of the precise synchronization of biological circadian rhythms with diurnal geomagnetic variation, an interplay involving the cryptochrome, MagR, and melatonin molecules.
